Specifgic Gene or Metabolite Levels

### 2025-01-14 Shortcuts_OmicsQuickPlotFunctions.Rmd
finalized omics functions

rawdata = read.csv(paste0(git.dir, "MetabolomicsForMetaboanalyst.txt"), sep = "\t", row.names = 1)

KEGG.Key = read.csv(paste0(git.dir, "RawMetabolomicData.csv"), row.names = 1)

KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

mean.metab = mean.data

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.metab[,condition1]/mean.metab[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.metab)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")[row.names(GRxWT.C),]
GRxWT.OE = compare.conditions("F", "A")[row.names(GRxWT.C),]

GR.CxDf = compare.conditions("D", "E")[row.names(GRxWT.C),]
WT.CxDf = compare.conditions("B", "C")[row.names(GRxWT.C),]

GR.CxOE = compare.conditions("F", "E")[row.names(GRxWT.C),]
WT.CxOE = compare.conditions("A", "C")[row.names(GRxWT.C),]

GR.DfxWT.C = compare.conditions("D", "C")[row.names(GRxWT.C),]
GR.OExWT.C = compare.conditions("F", "C")[row.names(GRxWT.C),]

comparison.means = list(GRF.CxDf = c("TktDfGR.F","GR.F"), 
                 GRF.CxOE = c("TktOEGR.F","GR.F"),
                 WTF.CxDf = c("TktDfWT.F","WT.F"),
                 WTF.CxOE = c("TktOEWT.F","WT.F"),
                 GRxWT.FC = c("GR.F","WT.F"), 
                 GRxWT.FDf = c("TktDfGR.F","TktDfWT.F"),
                 GRxWT.FOE = c("TktOEGR.F","TktOEWT.F"), 
                 GRM.CxDf = c("TktDfGR.M","GR.M"), 
                 GRM.CxOE = c("TktOEGR.M","GR.M"),
                 WTM.CxDf = c("TktDfWT.M","WT.M"),
                 WTM.CxOE = c("TktOEWT.M","WT.M"),
                 GRxWT.MC = c("GR.M","WT.M"), 
                 GRxWT.MDf = c("TktDfGR.M","TktDfWT.M"),
                 GRxWT.MOE = c("TktOEGR.M","TktOEWT.M"),
                 GR.FxM = c("GR.F","GR.M"),
                 WT.FxM = c("WT.F","WT.M"))

metab.FC = data.frame('GRxWT.C' = GRxWT.C$FC,
                      'GRxWT.Df' = GRxWT.Df$FC,
                      'GRxWT.OE' = GRxWT.OE$FC,
                      'GR.CxDf' = GR.CxDf$FC,
                      'WT.CxDf' = WT.CxDf$FC,
                      'GR.CxOE' = GR.CxOE$FC,
                      'WT.CxOE' = WT.CxOE$FC,
                      'GR.DfxWT.C' = GR.DfxWT.C$FC,
                      'GR.OExWT.C' = GR.OExWT.C$FC)
row.names(metab.FC) = row.names(GRxWT.C)
metab.FC = log2(metab.FC)

metab.FDR = data.frame(GRxWT.C = GRxWT.C$p,
                       GRxWT.Df = GRxWT.Df$p,
                       GRxWT.OE = GRxWT.OE$p,
                       GR.CxDf = GR.CxDf$p,
                       WT.CxDf = WT.CxDf$p,
                       GR.CxOE = GR.CxOE$p,
                       WT.CxOE = WT.CxOE$p,
                       GR.DfxWT.C = GR.DfxWT.C$p,
                       GR.OExWT.C = GR.OExWT.C$p)
row.names(metab.FDR) = row.names(GRxWT.C)

colnames(mean.metab) = as.character(genotypes[colnames(mean.metab)])

colnames(norm.data) = as.character(genotypes[substr(colnames(norm.data),1,1)])

@@

```{r echo = F}
##Function to plot all G85R groups
plot.G85R = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

@@

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$cpm), max(gene$cpm)),
     xlim = c(.5, 12.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

@@

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)
}

##Function to plot a specific gene expression levels in all A4V groups
plot.A4V = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
  FBgn = ID.row$FBgn
  name = ID.row$Symbol
  A4V.groups = as.vector(colnames(A4V.cpm))

@@

  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]),
                    condition = A4V.groups,
                    group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, start = nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

  gene$condition = factor(gene$condition, levels = c("A3FH", "A9FH", "A40FH",
                                                   "S3FH", "S9FH", "S40FH",
                                                   "A3FT", "A9FT", "A40FT",
                                                   "S3FT", "S9FT", "S40FT",
                                                   "A3FA", "A9FA", "A40FA",
                                                   "S3FA", "S9FA", "S40FA",
                                                   "A3MH", "A9MH",
                                                   "S3MH", "S9MH",
                                                   "A3MT", "A9MT",
                                                   "S3MT", "S9MT",
                                                   "A3MA", "A9MA",
                                                   "S3MA", "S9MA"))
  gene$group = factor(gene$group, levels = c(unique(gene$group)))

  boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
  points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
  legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)
}

@@

if(length(ID.row) < 50 & length(ID.row) > 0){
if(length(ID.row) == 1){
metab = row.names(norm.data)[ID.row]
name = row.names(norm.data)[ID.row]

@@

if(length(ID.row) != 1){
  ID.row
}

@@

##Function to plot a specific gene expression levels in all cell types for the SCdata
plot.SC = function(gene.name){
ID.row = SCdata[grep(gene.name, SCdata$Gene.name),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){

  gene.name = ID.row$Gene.name
  gene.levels = setNames(as.numeric(ID.row[,colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  gene.levels = gene.levels[18:1]
  highlight = rep(0, 18)
  highlight[(gene.levels > 2*(mean(gene.levels)))] = 1
  highlight = gene.levels * highlight

@@

ID = "Arc1"

ID = "Arc1"

metabID = 'NADH'

gene.name = 'gbb'

```{r echo = F}
## Plot a subset of groups in the TKT dataset

gene = data.frame(cpm = as.numeric(TKT.cpm[FBgn,]), 
                  condition = TKT.groups$Group, 
                  group = paste0(TKT.groups$Genotype, TKT.groups$Sex),
                  TKT = TKT.groups$TKT)

gene = gene[gene$TKT != 'OE',]

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('GR.F', "TktDfGR.F",
                                                 'WT.F', "TktDfWT.F",
                                                 'GR.M', "TktDfGR.M",
                                                 'WT.M', "TktDfWT.M"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

if(nrow(ID.row) != 1){
  ID.row
}
}

@@

if(nrow(ID.row) != 1){
  ID.row
}
}

@@

if(nrow(ID.row) != 1){
  ID.row
}
}

@@

>>>>

Metab.data = read.csv(paste0(git.dir, "RawMetabolomicData.csv"), row.names = 1)
Metab.Meta = read.csv(paste0(git.dir, "MetabolomicMetadata.csv"), row.names = 1)

KEGG.Key = setNames(Metab.data[-c(1:2), 2], row.names(Metab.data)[-c(1:2)])

norm.func = function(data){
  nd = (as.numeric(data)/Metab.Meta$TIC)*1000
  return(nd)
}

norm.data = t(apply(Metab.data[3:nrow(Metab.data),3:ncol(Metab.data)], 1, norm.func))

colnames(norm.data) = colnames(Metab.data)[3:ncol(Metab.data)]
colnames(norm.data) = Metab.Meta[colnames(norm.data), "Genotypes"]

@@

```{r echo = F}
##Function to plot all G85R groups
plot.G85R = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]
ID.row = na.omit(ID.row)

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1
   | length(intersect(GeneIDKey$FBgn, ID)) == 1
   | length(intersect(GeneIDKey$Symbol, ID)) == 1){
  if(nrow(ID.row) == 1){
    FBgn = ID.row$FBgn
    name = ID.row$Symbol
  }
  if(length(intersect(GeneIDKey$FBgn, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$FBgn == ID]
    name = ID.row$Symbol[ID.row$FBgn == ID]
  }
  if(length(intersect(GeneIDKey$Symbol, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$Symbol == ID]
    name = ID.row$Symbol[ID.row$Symbol == ID]
  }

@@

if(nrow(na.omit(gene)) > 0){
par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$cpm), max(gene$cpm)),
     xlim = c(.5, 12.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

@@

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)
}
}
}

if(nrow(ID.row) != 1
   | length(intersect(GeneIDKey$FBgn, ID)) > 1
   | length(intersect(GeneIDKey$Symbol, ID)) > 1){
  ID.row
}
}

##Function to plot a specific gene expression levels in all A4V groups
plot.A4V = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]
ID.row = na.omit(ID.row)

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1
   | length(intersect(GeneIDKey$FBgn, ID)) == 1
   | length(intersect(GeneIDKey$Symbol, ID)) == 1){
  if(nrow(ID.row) == 1){
    FBgn = ID.row$FBgn
    name = ID.row$Symbol
  }
  if(length(intersect(GeneIDKey$FBgn, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$FBgn == ID]
    name = ID.row$Symbol[ID.row$FBgn == ID]
  }
  if(length(intersect(GeneIDKey$Symbol, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$Symbol == ID]
    name = ID.row$Symbol[ID.row$Symbol == ID]
  }
  A4V.groups = as.vector(colnames(A4V.cpm))

@@

  if(nrow(na.omit(A4V.cpm[FBgn,]))){
  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]),
                    condition = A4V.groups,
                    group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, start = nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

  gene$condition = factor(gene$condition, levels = c("S3FH", "S9FH", "S40FH",
                                                     "S3FT", "S9FT", "S40FT",
                                                     "S3FA", "S9FA", "S40FA",
                                                     "A3FH", "A9FH", "A40FH",
                                                     "A3FT", "A9FT", "A40FT",
                                                     "A3FA", "A9FA", "A40FA",
                                                     "S3MH", "S9MH",
                                                     "A3MT", "A9MT",
                                                     "A3MA", "A9MA",
                                                     "A3MH", "A9MH",
                                                     "S3MT", "S9MT",
                                                     "S3MA", "S9MA"))
  gene$group = factor(gene$group, levels = c("SF", "AF", "SM", "AM"))

  boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
  points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
  legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent M'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)
  }
}

if(nrow(ID.row) != 1
   | length(intersect(GeneIDKey$FBgn, ID)) > 1
   | length(intersect(GeneIDKey$Symbol, ID)) > 1){
  ID.row
}
}
}

@@

if(length(ID.row) < 50 & length(ID.row) > 0){
if(length(ID.row) == 1 | length(intersect(row.names(norm.data), metabID)) == 1){

  if(length(ID.row) == 1){
    metab = row.names(norm.data)[ID.row]
    name = row.names(norm.data)[ID.row]
  }

  if(length(intersect(row.names(norm.data), metabID)) == 1){
    metab = metabID
    name = metabID
  }

@@

if(length(ID.row) != 1 | length(intersect(row.names(norm.data), metabID)) > 1){
  norm.data[ID.row,]
}

@@

##Function to plot a specific gene expression levels in all cell types for the SC data
plot.SC = function(gene.name){
ID.row = SCdata[grep(gene.name, SCdata$Gene.name),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1 | length(intersect(gene.name, SCdata$Gene.name)) == 1){

  if(length(intersect(gene.name, SCdata$Gene.name)) == 1){
    gene.levels = setNames(as.numeric(ID.row[ID.row$Gene.name == gene.name, colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  }

  if(nrow(ID.row) == 1){
    gene.name = ID.row$Gene.name
    gene.levels = setNames(as.numeric(ID.row[,colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  }

  gene.levels = gene.levels[18:1]
  highlight = rep(0, 18)
  highlight[(gene.levels > 2*(mean(gene.levels)))] = 1
  highlight = gene.levels * highlight

@@

if(nrow(ID.row) != 1 | length(intersect(gene.name, SCdata$Gene.name)) > 1){
  ID.row
}
}
}

##Function to plot a specific gene expression levels in the gbbOE transcriptomic data set
plot.gbbOE = function(ID){

@@

##Function to plot a specific gene expression levels in the gbbKO transcriptomic data set
plot.gbbKO = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

@@

@@

ID = "ETH"

ID = "ETH"

metabID = 'NAD'

gene.name = 'CCAP'

plot.gbbOE(ID)

plot.gbbKO(ID)
